---
title: "SPiCT"
output: 
  html_document:
    theme: cerulean
    highlight: pygments 
    fig_caption: false
    df_print: kable
    toc: true
    toc_float: 
      collapsed: true
    number_sections: false
    self_contained: true
    code_download: true
---
<style> body {text-align: justify} </style> <!-- Justify text. -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

```{r, include = FALSE}
# Paquetes ---------------------------
library(spict)
library(knitr)
library(formatR)
library(ellipse)
library(ggplot2)
library(corrplot)
library(icesAdvice)
library(Metrics)
library(MLmetrics)

# Load data
load("./simulated_data.RData")
```

# Relativa biomass indices (random sampling)

## Inputs SCiPT {.tabset .tabset-fade}

```{r, eval = FALSE}
# INPUT 1: catch ---------------------------
## Extract the simulated biomass each year ---------------------------
simulated_biomass <- list()
for (i in 1:k) {
  simulated_biomass[[i]] <- sum(data_variables[[i]]$biomasa)
}
simulated_biomass <- unlist(simulated_biomass)
## Set parameters of Schafer's production curve ---------------------------
r <- 1.4
K <- sum(simulated_biomass)/0.5

## Catch approximation ---------------------------
simulated_catch <- list()
for (i in 1: (k-1)) {
  simulated_catch[[i]] <- simulated_biomass[i] - simulated_biomass[(i + 1)] + (r*simulated_biomass[i] * (1-(simulated_biomass[i]/K)))
}

simulated_catch[[15]] <-  r*simulated_biomass[15]*(1-(simulated_biomass[15]/K)) # We assumed equilibrium biomass_14 = biomass_15 

simulated_catch <- unlist(simulated_catch)

## Capturas ---------------------------
C_sp_i <- data.frame(obsC = simulated_catch, timeC = rep(2000:2014)) 

# INPUT 2: relative indices of biomass ---------------------------
load("./random_series.RData")
I_sp_i <- data.frame(ObsI = modelo, timeI = 2000:2014) 

# Joint INPUTS ---------------------------
inp <- list(
  timeC = C_sp_i$timeC, # Year
  obsC = C_sp_i$obsC, # Catch
  timeI = list(I_sp_i$timeI), # Year
  obsI = list(I_sp_i$ObsI) # relativa indices
)

inp <- check.inp(inp)
inp$dtc
inp$dteuler <- 1/16
inp$ini$logn <- log(2) # Set p
inp$phases$logn <- -1
```

```{r, include = FALSE}
# INPUT 1: catch ---------------------------
## Extract the simulated biomass each year ---------------------------
simulated_biomass <- list()
for (i in 1:k) {
  simulated_biomass[[i]] <- sum(data_variables[[i]]$biomasa)
}
simulated_biomass <- unlist(simulated_biomass)
## Set parameters of Schafer's production curve ---------------------------
r <- 1.4
K <- sum(simulated_biomass)/0.5

## Catch approximation ---------------------------
simulated_catch <- list()
for (i in 1: (k-1)) {
  simulated_catch[[i]] <- simulated_biomass[i] - simulated_biomass[(i + 1)] + (r*simulated_biomass[i] * (1-(simulated_biomass[i]/K)))
}

simulated_catch[[15]] <-  r*simulated_biomass[15]*(1-(simulated_biomass[15]/K)) # We assumed equilibrium biomass_14 = biomass_15 

simulated_catch <- unlist(simulated_catch)

## Capturas ---------------------------
C_sp_i <- data.frame(obsC = simulated_catch, timeC = rep(2000:2014)) 

# INPUT 2: relative indices of biomass ---------------------------
load("./random_series.RData")
I_sp_i_1 <- data.frame(ObsI = glm_f_bio_r, timeI = 2000:2014) 
I_sp_i_2 <- data.frame(ObsI = glm_b_bio_r, timeI = 2000:2014) 
I_sp_i_3 <- data.frame(ObsI = gam_f_bio_r, timeI = 2000:2014) 
I_sp_i_4 <- data.frame(ObsI = gam_b_bio_r, timeI = 2000:2014) 
I_sp_i_5 <- data.frame(ObsI = geo_b_bio_r, timeI = 2000:2014) 

# Joint INPUTS ---------------------------
## GLM F
inp_1 <- list(
  timeC = C_sp_i$timeC, # Year
  obsC = C_sp_i$obsC, # Catch
  timeI = list(I_sp_i_1$timeI), # Year
  obsI = list(I_sp_i_1$ObsI) # relative indices
)

inp_1 <- check.inp(inp_1)
inp_1$dtc
inp_1$dteuler <- 1/16
inp_1$ini$logn <- log(2) # Set p
inp_1$phases$logn <- -1

## GLM B 
inp_2 <- list(
  timeC = C_sp_i$timeC, # Year
  obsC = C_sp_i$obsC, # Catch
  timeI = list(I_sp_i_2$timeI), # Year
  obsI = list(I_sp_i_2$ObsI) # relative indices
)

inp_2 <- check.inp(inp_2)
inp_2$dtc
inp_2$dteuler <- 1/16
inp_2$ini$logn <- log(2) # Set p
inp_2$phases$logn <- -1

## GAM F
inp_3 <- list(
  timeC = C_sp_i$timeC, # Year
  obsC = C_sp_i$obsC, # Catch
  timeI = list(I_sp_i_3$timeI), # Year
  obsI = list(I_sp_i_3$ObsI) # relative indices
)

inp_3 <- check.inp(inp_3)
inp_3$dtc
inp_3$dteuler <- 1/16
inp_3$ini$logn <- log(2) # Set p
inp_3$phases$logn <- -1

## GAM B
inp_4 <- list(
  timeC = C_sp_i$timeC, # Year
  obsC = C_sp_i$obsC, # Catch
  timeI = list(I_sp_i_4$timeI), # Year
  obsI = list(I_sp_i_4$ObsI) # relative indices
)

inp_4 <- check.inp(inp_4)
inp_4$dtc
inp_4$dteuler <- 1/16
inp_4$ini$logn <- log(2) # Set p
inp_4$phases$logn <- -1

## GEO
inp_5 <- list(
  timeC = C_sp_i$timeC, # Year
  obsC = C_sp_i$obsC, # Catch
  timeI = list(I_sp_i_5$timeI), # Year
  obsI = list(I_sp_i_5$ObsI) # relative indices
)

inp_5 <- check.inp(inp_5)
inp_5$dtc
inp_5$dteuler <- 1/16
inp_5$ini$logn <- log(2) # Set p
inp_5$phases$logn <- -1
```

### GLM (frequentist)

```{r, echo = FALSE, fig.align='center'}
plotspict.data(inp_1)
```

### GLM (Bayesian)

```{r, echo = FALSE, fig.align='center'}
plotspict.data(inp_2)
```

### GAM (frequentist)

```{r, echo = FALSE, fig.align='center'}
plotspict.data(inp_3)
```

### GAM (Bayesian)

```{r, echo = FALSE, fig.align='center'}
plotspict.data(inp_4)
```

### Geostatistic

```{r, echo = FALSE, fig.align='center'}
plotspict.data(inp_5)
```

## Fit the model {.tabset .tabset-fade}

```{r, eval = FALSE}
# SPiCT ---------------------------
res <- fit.spict(inp)
```

```{r, include = FALSE}
res_1 <- fit.spict(inp_1) # GLM F
res_2 <- fit.spict(inp_2) # GLM B
res_3 <- fit.spict(inp_3) # GAM F
res_4 <- fit.spict(inp_4) # GAM B
res_5 <- fit.spict(inp_5) # GEO
```

### GLM (frequentist)

```{r, echo = FALSE, fig.align='center'}
plot(res_1)
```

### GLM (Bayesian)

```{r, echo = FALSE, fig.align='center'}
plot(res_2)
```

### GAM (frequentist)

```{r, echo = FALSE, fig.align='center'}
plot(res_3)
```

### GAM (Bayesian)

```{r, echo = FALSE, fig.align='center'}
plot(res_4)
```

### Geostatistic

```{r, echo = FALSE, fig.align='center'}
plot(res_5)
```

## Diagnostic {.tabset .tabset-fade}

```{r, eval = FALSE}
# Diagnostic ---------------------------
## Convergence 
res$opt$convergence # 0 convergence  
all(is.finite(res$sd)) # if it is TRUE is okay 

residuals <- calc.osa.resid(res) # residuals of the model
plotspict.diagnostic(residuals) # test for the diagnostic

## Retrospective analysis
r <- fit.spict(inp) 
rep=retro(r, nretroyear=3) # Eliminate 3 years
plotspict.retro(rep) # plot for the retrospective analysis 
m1=mohns_rho(rep, what = c("FFmsy", "BBmsy"));
m1 # Discrepancy between last year and trajectories, it is NOT suitable for our specific case


## Production curve
plotspict.production(res) # check that the points fall around the curve

## Model robustness (change of initial values)
set.seed(1234)
a=check.ini(inp, ntrials=30 
            ,verbose = FALSE)
a$check.ini$propchng
a$check.ini$resmat # check distance is close to 0 

# Confidence intervals
fit=res # no more than 1 unit
get.par("logBmBmsy", fit, exp=TRUE, CI = 0.9) 
get.par("logFmFmsy", fit, exp=TRUE, CI = 0.9)

# Estimated biomass vs real biomass
get.par("logB", fit, exp=TRUE, CI = 0.9) 
biomass_scpit <- get.par("logB", res, exp = TRUE)[seq(1,length(inp$time),by=16),2]
biomass_scpit <- biomass_scpit[1:15] 
plot(biomass_scpit, col = "red", type = "l")
plot(simulated_biomass, col = "blue", type = "l")
plotspict.biomass(res) 
```

### GLM (frequentist)

#### Measures for convergence

```{r}
# Diagnostic ---------------------------
## Convergence 
res_1$opt$convergence # 0 convergence  
all(is.finite(res_1$sd)) # if it is TRUE is okay 
```

#### Residuals analysis

```{r, fig.align='center'}
residuals_1 <- calc.osa.resid(res_1) # residuals of the model
plotspict.diagnostic(residuals_1) # test for the diagnostic
```

#### Retrospective analysis

```{r, fig.align='center'}
## Retrospective analysis
r_1 <- fit.spict(inp_1) 
rep_1 = retro(r_1, nretroyear=3) # Eliminate 3 years
plotspict.retro(rep_1) # plot for the retrospective analysis 
```

#### Production curve

```{r, fig.align='center'}
## Production curve
plotspict.production(res_1) # check that the points fall around the curve
```

#### Model robustness

```{r}
## Model robustness (change of initial values)
set.seed(1234)
a_1=check.ini(inp_1, ntrials=30 
            ,verbose = FALSE)
#a_1$check.ini$propchng
table(a_1$check.ini$resmat[1:30] == 0) # check distance is close to 0 
```

#### Confidence intervals

```{r}
# Confidence intervals
fit_1=res_1 # no more than 1 unit
get.par("logBmBmsy", fit_1, exp=TRUE, CI = 0.9) 
get.par("logFmFmsy", fit_1, exp=TRUE, CI = 0.9)
```

#### Estimated biomass vs Simulated biomass

```{r}
# get.par("logB", fit_1, exp=TRUE, CI = 0.9) 
biomass_SPiCT_1 <- get.par("logB", res_1, exp = TRUE)[seq(1,length(inp_1$time),by=16),2]
biomass_SPiCT_1 <- biomass_SPiCT_1[1:15] 
```

```{r, fig.align= TRUE}
plot(biomass_SPiCT_1,
     col = "red",
     type = "l",
     ylim = c(min(simulated_biomass), max(biomass_SPiCT_1))
     )
lines(simulated_biomass, col = "blue", type = "l")
```

```{r}
# Error measures
glm_f_rmse <- rmse(biomass_SPiCT_1, simulated_biomass)
glm_f_mae <- mae(biomass_SPiCT_1, simulated_biomass)
glm_f_mape <- MAPE(biomass_SPiCT_1, simulated_biomass)
glm_f_rmse # RMSE
glm_f_mae # MAE
glm_f_mape # MAPE
```

### GLM (Bayesian)

#### Measures for convergence

```{r}
# Diagnostic ---------------------------
## Convergence 
res_2$opt$convergence # 0 convergence  
all(is.finite(res_2$sd)) # if it is TRUE is okay 
```

#### Residuals analysis

```{r, fig.align='center'}
residuals_2 <- calc.osa.resid(res_2) # residuals of the model
plotspict.diagnostic(residuals_2) # test for the diagnostic
```

#### Retrospective analysis

```{r, fig.align='center'}
## Retrospective analysis
r_2 <- fit.spict(inp_2) 
rep_2 = retro(r_2, nretroyear=3) # Eliminate 3 years
plotspict.retro(rep_2) # plot for the retrospective analysis 
```

#### Production curve

```{r, fig.align='center'}
## Production curve
plotspict.production(res_2) # check that the points fall around the curve
```

#### Model robustness

```{r}
## Model robustness (change of initial values)
set.seed(1234)
a_2=check.ini(inp_2, ntrials=30 
            ,verbose = FALSE)
#a_2$check.ini$propchng
table(a_2$check.ini$resmat[1:30] == 0) # check distance is close to 0 
```

#### Confidence intervals

```{r}
# Confidence intervals
fit_2=res_2 # no more than 1 unit
get.par("logBmBmsy", fit_2, exp=TRUE, CI = 0.9) 
get.par("logFmFmsy", fit_2, exp=TRUE, CI = 0.9)
```

#### Estimated biomass vs Simulated biomass

```{r}
# get.par("logB", fit_2, exp=TRUE, CI = 0.9) 
biomass_SPiCT_2 <- get.par("logB", res_2, exp = TRUE)[seq(1,length(inp_2$time),by=16),2]
biomass_SPiCT_2 <- biomass_SPiCT_2[1:15] 
```

```{r, fig.align= TRUE}
plot(biomass_SPiCT_2,
     col = "red",
     type = "l",
     ylim = c(min(simulated_biomass), max(biomass_SPiCT_2))
     )
lines(simulated_biomass, col = "blue", type = "l")
```

```{r}
# Error measures
glm_b_rmse <- rmse(biomass_SPiCT_2, simulated_biomass)
glm_b_mae <- mae(biomass_SPiCT_2, simulated_biomass)
glm_b_mape <- MAPE(biomass_SPiCT_2, simulated_biomass)
glm_b_rmse # RMSE
glm_b_mae # MAE
glm_b_mape # MAPE
```

### GAM (frequentist)

#### Measures for convergence

```{r}
# Diagnostic ---------------------------
## Convergence 
res_3$opt$convergence # 0 convergence  
all(is.finite(res_3$sd)) # if it is TRUE is okay 
```

#### Residuals analysis

```{r, fig.align='center'}
residuals_3 <- calc.osa.resid(res_3) # residuals of the model
plotspict.diagnostic(residuals_3) # test for the diagnostic
```

#### Retrospective analysis

```{r, fig.align='center'}
## Retrospective analysis
r_3 <- fit.spict(inp_3) 
rep_3 = retro(r_3, nretroyear=3) # Eliminate 3 years
plotspict.retro(rep_3) # plot for the retrospective analysis 
```

#### Production curve

```{r, fig.align='center'}
## Production curve
plotspict.production(res_3) # check that the points fall around the curve
```

#### Model robustness

```{r}
## Model robustness (change of initial values)
set.seed(1234)
a_3=check.ini(inp_3, ntrials=30 
            ,verbose = FALSE)
#a_3$check.ini$propchng
table(a_3$check.ini$resmat[1:30] == 0) # check distance is close to 0 
```

#### Confidence intervals

```{r}
# Confidence intervals
fit_3=res_3 # no more than 1 unit
get.par("logBmBmsy", fit_3, exp=TRUE, CI = 0.9) 
get.par("logFmFmsy", fit_3, exp=TRUE, CI = 0.9)
```

#### Estimated biomass vs Simulated biomass

```{r}
# get.par("logB", fit_3, exp=TRUE, CI = 0.9) 
biomass_SPiCT_3 <- get.par("logB", res_3, exp = TRUE)[seq(1,length(inp_3$time),by=16),2]
biomass_SPiCT_3 <- biomass_SPiCT_3[1:15] 
```

```{r, fig.align= TRUE}
plot(biomass_SPiCT_3,
     col = "red",
     type = "l",
     ylim = c(min(simulated_biomass), max(biomass_SPiCT_3))
     )
lines(simulated_biomass, col = "blue", type = "l")
```

```{r}
# Error measures
gam_f_rmse <- rmse(biomass_SPiCT_3, simulated_biomass)
gam_f_mae <- mae(biomass_SPiCT_3, simulated_biomass)
gam_f_mape <- MAPE(biomass_SPiCT_3, simulated_biomass)
gam_f_rmse # RMSE
gam_f_mae # MAE
gam_f_mape # MAPE
```

### GAM (Bayesian)

#### Measures for convergence

```{r}
# Diagnostic ---------------------------
## Convergence 
res_4$opt$convergence # 0 convergence  
all(is.finite(res_4$sd)) # if it is TRUE is okay 
```

#### Residuals analysis

```{r, fig.align='center'}
residuals_4 <- calc.osa.resid(res_4) # residuals of the model
plotspict.diagnostic(residuals_4) # test for the diagnostic
```

#### Retrospective analysis

```{r, fig.align='center'}
## Retrospective analysis
r_4 <- fit.spict(inp_4) 
rep_4 = retro(r_4, nretroyear=3) # Eliminate 3 years
plotspict.retro(rep_4) # plot for the retrospective analysis 
```

#### Production curve

```{r, fig.align='center'}
## Production curve
plotspict.production(res_4) # check that the points fall around the curve
```

#### Model robustness

```{r}
## Model robustness (change of initial values)
set.seed(1234)
a_4=check.ini(inp_4, ntrials=30 
            ,verbose = FALSE)
#a_4$check.ini$propchng
table(a_4$check.ini$resmat[1:30] == 0) # check distance is close to 0 
```

#### Confidence intervals

```{r}
# Confidence intervals
fit_4=res_4 # no more than 1 unit
get.par("logBmBmsy", fit_4, exp=TRUE, CI = 0.9) 
get.par("logFmFmsy", fit_4, exp=TRUE, CI = 0.9)
```

#### Estimated biomass vs Simulated biomass

```{r}
# get.par("logB", fit_4, exp=TRUE, CI = 0.9) 
biomass_SPiCT_4 <- get.par("logB", res_4, exp = TRUE)[seq(1,length(inp_4$time),by=16),2]
biomass_SPiCT_4 <- biomass_SPiCT_4[1:15] 
```

```{r, fig.align= TRUE}
plot(biomass_SPiCT_4,
     col = "red",
     type = "l",
     ylim = c(min(simulated_biomass), max(biomass_SPiCT_4))
     )
lines(simulated_biomass, col = "blue", type = "l")
```

```{r}
# Error measures
gam_b_rmse <- rmse(biomass_SPiCT_4, simulated_biomass)
gam_b_mae <- mae(biomass_SPiCT_4, simulated_biomass)
gam_b_mape <- MAPE(biomass_SPiCT_4, simulated_biomass)
gam_b_rmse # RMSE
gam_b_mae # MAE
gam_b_mape # MAPE
```

### Geostatistic

#### Measures for convergence

```{r}
# Diagnostic ---------------------------
## Convergence 
res_5$opt$convergence # 0 convergence  
all(is.finite(res_5$sd)) # if it is TRUE is okay 
```

#### Residuals analysis

```{r, fig.align='center'}
residuals_5 <- calc.osa.resid(res_5) # residuals of the model
plotspict.diagnostic(residuals_5) # test for the diagnostic
```

#### Retrospective analysis

```{r, fig.align='center'}
## Retrospective analysis
r_5 <- fit.spict(inp_5) 
rep_5 = retro(r_5, nretroyear=3) # Eliminate 3 years
plotspict.retro(rep_5) # plot for the retrospective analysis 
```

#### Production curve

```{r, fig.align='center'}
## Production curve
plotspict.production(res_5) # check that the points fall around the curve
```

#### Model robustness

```{r}
## Model robustness (change of initial values)
set.seed(1234)
a_5=check.ini(inp_5, ntrials=30 
            ,verbose = FALSE)
#a_5$check.ini$propchng
table(a_5$check.ini$resmat[1:30] == 0) # check distance is close to 0 
```

#### Confidence intervals

```{r}
# Confidence intervals
fit_5=res_5 # no more than 1 unit
get.par("logBmBmsy", fit_5, exp=TRUE, CI = 0.9) 
get.par("logFmFmsy", fit_5, exp=TRUE, CI = 0.9)
```

#### Estimated biomass vs Simulated biomass

```{r}
# get.par("logB", fit_5, exp=TRUE, CI = 0.9) 
biomass_SPiCT_5 <- get.par("logB", res_5, exp = TRUE)[seq(1,length(inp_5$time),by=16),2]
biomass_SPiCT_5 <- biomass_SPiCT_5[1:15] 
```

```{r, fig.align= TRUE}
plot(biomass_SPiCT_5,
     col = "red",
     type = "l",
     ylim = c(min(simulated_biomass), max(biomass_SPiCT_5))
     )
lines(simulated_biomass, col = "blue", type = "l")
```

```{r}
# Error measures
geo_b_rmse <- rmse(biomass_SPiCT_5, simulated_biomass)
geo_b_mae <- mae(biomass_SPiCT_5, simulated_biomass)
geo_b_mape <- MAPE(biomass_SPiCT_5, simulated_biomass)
geo_b_rmse # RMSE
geo_b_mae # MAE
geo_b_mape # MAPE
```

```{r, include = FALSE}
data_error_random <- data.frame(models = c("GLM F",
                                           "GLM B",
                                           "GAM F",
                                           "GAM B",
                                           "GEO"),
  
                              rmse = c(glm_f_rmse,
                                        glm_b_rmse,
                                        gam_f_rmse,
                                        gam_b_rmse,
                                        geo_b_rmse),
                               
                                mae = c(glm_f_mae,
                                        glm_b_mae,
                                        gam_f_mae,
                                        gam_b_mae,
                                        geo_b_mae),
                               
                                mape = c(glm_f_mape,
                                         glm_b_mape,
                                         gam_f_mape,
                                         gam_b_mape,
                                         geo_b_mape)
                               )

```

# CPUE indices (preferential sampling)

## INPUTS SPiCT {.tabset .tabset-fade}

```{r, include = FALSE}
# INPUT 1: catch ---------------------------
## Extract the simulated biomass each year ---------------------------
simulated_biomass <- list()
for (i in 1:k) {
  simulated_biomass[[i]] <- sum(data_variables[[i]]$biomasa)
}
simulated_biomass <- unlist(simulated_biomass)
## Set parameters of Schafer's production curve ---------------------------
r <- 1.4
K <- sum(simulated_biomass)/0.5

## Catch approximation ---------------------------
simulated_catch <- list()
for (i in 1: (k-1)) {
  simulated_catch[[i]] <- simulated_biomass[i] - simulated_biomass[(i + 1)] + (r*simulated_biomass[i] * (1-(simulated_biomass[i]/K)))
}

simulated_catch[[15]] <-  r*simulated_biomass[15]*(1-(simulated_biomass[15]/K)) # We assumed equilibrium biomass_14 = biomass_15 

simulated_catch <- unlist(simulated_catch)

## Capturas ---------------------------
C_sp_i <- data.frame(obsC = simulated_catch, timeC = rep(2000:2014)) 

# INPUT 2: relative indices of biomass ---------------------------
load("./pref_series.RData")
I_sp_i_1 <- data.frame(ObsI = glm_f_CPUE, timeI = 2000:2014) 
I_sp_i_2 <- data.frame(ObsI = glm_b_CPUE, timeI = 2000:2014) 
I_sp_i_3 <- data.frame(ObsI = gam_f_CPUE, timeI = 2000:2014) 
I_sp_i_4 <- data.frame(ObsI = gam_b_CPUE, timeI = 2000:2014) 
I_sp_i_5 <- data.frame(ObsI = geo_b_CPUE, timeI = 2000:2014)
I_sp_i_6 <- data.frame(ObsI = geo_pref_b_CPUE, timeI = 2000:2014) 


# Joint INPUTS ---------------------------
## GLM F
inp_1 <- list(
  timeC = C_sp_i$timeC, # Year
  obsC = C_sp_i$obsC, # Catch
  timeI = list(I_sp_i_1$timeI), # Year
  obsI = list(I_sp_i_1$ObsI) # relative indices
)

inp_1 <- check.inp(inp_1)
inp_1$dtc
inp_1$dteuler <- 1/16
inp_1$ini$logn <- log(2) # Set p
inp_1$phases$logn <- -1

## GLM B 
inp_2 <- list(
  timeC = C_sp_i$timeC, # Year
  obsC = C_sp_i$obsC, # Catch
  timeI = list(I_sp_i_2$timeI), # Year
  obsI = list(I_sp_i_2$ObsI) # relative indices
)

inp_2 <- check.inp(inp_2)
inp_2$dtc
inp_2$dteuler <- 1/16
inp_2$ini$logn <- log(2) # Set p
inp_2$phases$logn <- -1

## GAM F
inp_3 <- list(
  timeC = C_sp_i$timeC, # Year
  obsC = C_sp_i$obsC, # Catch
  timeI = list(I_sp_i_3$timeI), # Year
  obsI = list(I_sp_i_3$ObsI) # relative indices
)

inp_3 <- check.inp(inp_3)
inp_3$dtc
inp_3$dteuler <- 1/16
inp_3$ini$logn <- log(2) # Set p
inp_3$phases$logn <- -1

## GAM B
inp_4 <- list(
  timeC = C_sp_i$timeC, # Year
  obsC = C_sp_i$obsC, # Catch
  timeI = list(I_sp_i_4$timeI), # Year
  obsI = list(I_sp_i_4$ObsI) # relative indices
)

inp_4 <- check.inp(inp_4)
inp_4$dtc
inp_4$dteuler <- 1/16
inp_4$ini$logn <- log(2) # Set p
inp_4$phases$logn <- -1

## GEO 
inp_5 <- list(
  timeC = C_sp_i$timeC, # Year
  obsC = C_sp_i$obsC, # Catch
  timeI = list(I_sp_i_5$timeI), # Year
  obsI = list(I_sp_i_5$ObsI) # relative indices
)

inp_5 <- check.inp(inp_5)
inp_5$dtc
inp_5$dteuler <- 1/16
inp_5$ini$logn <- log(2) # Set p
inp_5$phases$logn <- -1

## MARKED POINT PATTERN 
inp_6 <- list(
  timeC = C_sp_i$timeC, # Year
  obsC = C_sp_i$obsC, # Catch
  timeI = list(I_sp_i_6$timeI), # Year
  obsI = list(I_sp_i_6$ObsI) # relative indices
)

inp_6 <- check.inp(inp_6)
inp_6$dtc
inp_6$dteuler <- 1/16
inp_6$ini$logn <- log(2) # Set p
inp_6$phases$logn <- -1
```

### GLM (frequentist)

```{r, echo = FALSE, fig.align='center'}
plotspict.data(inp_1)
```

### GLM (Bayesian)

```{r, echo = FALSE, fig.align='center'}
plotspict.data(inp_2)
```

### GAM (frequentist)

```{r, echo = FALSE, fig.align='center'}
plotspict.data(inp_3)
```

### GAM (Bayesian)

```{r, echo = FALSE, fig.align='center'}
plotspict.data(inp_4)
```

### Geostatistic

```{r, echo = FALSE, fig.align='center'}
plotspict.data(inp_5)
```

### Marked point pattern

```{r, echo = FALSE, fig.align='center'}
plotspict.data(inp_6)
```

## Fit the model {.tabset .tabset-fade}

```{r, eval = FALSE}
# SPiCT ---------------------------
res <- fit.spict(inp)
```

```{r, include = FALSE}
res_1 <- fit.spict(inp_1) # GLM F
res_2 <- fit.spict(inp_2) # GLM B
res_3 <- fit.spict(inp_3) # GAM F
res_4 <- fit.spict(inp_4) # GAM B
res_5 <- fit.spict(inp_5) # GEO
res_6 <- fit.spict(inp_6) # MPP
```

### GLM (frequentist)

```{r, echo = FALSE, fig.align='center'}
plot(res_1)
```

### GLM (Bayesian)

```{r, echo = FALSE, fig.align='center'}
plot(res_2)
```

### GAM (frequentist)

```{r, echo = FALSE, fig.align='center'}
plot(res_3)
```

### GAM (Bayesian)

```{r, echo = FALSE, fig.align='center'}
plot(res_4)
```

### Geostatistic

```{r, echo = FALSE, fig.align='center'}
plot(res_5)
```

### Marked point pattern

```{r, echo = FALSE, fig.align='center'}
plot(res_6)
```

## Diagnostic {.tabset .tabset-fade}

```{r, eval = FALSE}
# Diagnostic ---------------------------
## Convergence 
res$opt$convergence # 0 convergence  
all(is.finite(res$sd)) # if it is TRUE is okay 

residuals <- calc.osa.resid(res) # residuals of the model
plotspict.diagnostic(residuals) # test for the diagnostic

## Retrospective analysis
r <- fit.spict(inp) 
rep=retro(r, nretroyear=3) # Eliminate 3 years
plotspict.retro(rep) # plot for the retrospective analysis 
m1=mohns_rho(rep, what = c("FFmsy", "BBmsy"));
m1 # Discrepancy between last year and trajectories, it is NOT suitable for our specific case


## Production curve
plotspict.production(res) # check that the points fall around the curve

## Model robustness (change of initial values)
set.seed(1234)
a=check.ini(inp, ntrials=30 
            ,verbose = FALSE)
a$check.ini$propchng
a$check.ini$resmat # check distance is close to 0 

# Confidence intervals
fit=res # no more than 1 unit
get.par("logBmBmsy", fit, exp=TRUE, CI = 0.9) 
get.par("logFmFmsy", fit, exp=TRUE, CI = 0.9)

# Estimated biomass vs real biomass
get.par("logB", fit, exp=TRUE, CI = 0.9) 
biomass_scpit <- get.par("logB", res, exp = TRUE)[seq(1,length(inp$time),by=16),2]
biomass_scpit <- biomass_scpit[1:15] 
plot(biomass_scpit, col = "red", type = "l")
plot(simulated_biomass, col = "blue", type = "l")
plotspict.biomass(res) 
```

### GLM (frequentist)

#### Measures for convergence

```{r}
# Diagnostic ---------------------------
## Convergence 
res_1$opt$convergence # 0 convergence  
all(is.finite(res_1$sd)) # if it is TRUE is okay 
```

#### Residuals analysis

```{r, fig.align='center'}
residuals_1 <- calc.osa.resid(res_1) # residuals of the model
plotspict.diagnostic(residuals_1) # test for the diagnostic
```

#### Retrospective analysis

```{r, fig.align='center'}
## Retrospective analysis
r_1 <- fit.spict(inp_1) 
rep_1 = retro(r_1, nretroyear=3) # Eliminate 3 years
plotspict.retro(rep_1) # plot for the retrospective analysis 
```

#### Production curve

```{r, fig.align='center'}
## Production curve
plotspict.production(res_1) # check that the points fall around the curve
```

#### Model robustness

```{r}
## Model robustness (change of initial values)
set.seed(1234)
a_1=check.ini(inp_1, ntrials=30 
            ,verbose = FALSE)
#a_1$check.ini$propchng
table(a_1$check.ini$resmat[1:30] == 0) # check distance is close to 0 
```

#### Confidence intervals

```{r}
# Confidence intervals
fit_1=res_1 # no more than 1 unit
get.par("logBmBmsy", fit_1, exp=TRUE, CI = 0.9) 
get.par("logFmFmsy", fit_1, exp=TRUE, CI = 0.9)
```

#### Estimated biomass vs Simulated biomass

```{r}
# get.par("logB", fit_1, exp=TRUE, CI = 0.9) 
biomass_SPiCT_1 <- get.par("logB", res_1, exp = TRUE)[seq(1,length(inp_1$time),by=16),2]
biomass_SPiCT_1 <- biomass_SPiCT_1[1:15] 
```

```{r, fig.align= TRUE}
plot(biomass_SPiCT_1,
     col = "red",
     type = "l",
     ylim = c(min(simulated_biomass), max(biomass_SPiCT_1))
     )
lines(simulated_biomass, col = "blue", type = "l")
```

```{r}
# Error measures
glm_f_rmse <- rmse(biomass_SPiCT_1, simulated_biomass)
glm_f_mae <- mae(biomass_SPiCT_1, simulated_biomass)
glm_f_mape <- MAPE(biomass_SPiCT_1, simulated_biomass)
glm_f_rmse # RMSE
glm_f_mae # MAE
glm_f_mape # MAPE
```

### GLM (Bayesian)

#### Measures for convergence

```{r}
# Diagnostic ---------------------------
## Convergence 
res_2$opt$convergence # 0 convergence  
all(is.finite(res_2$sd)) # if it is TRUE is okay 
```

#### Residuals analysis

```{r, fig.align='center'}
residuals_2 <- calc.osa.resid(res_2) # residuals of the model
plotspict.diagnostic(residuals_2) # test for the diagnostic
```

#### Retrospective analysis

```{r, fig.align='center'}
## Retrospective analysis
r_2 <- fit.spict(inp_2) 
rep_2 = retro(r_2, nretroyear=3) # Eliminate 3 years
plotspict.retro(rep_2) # plot for the retrospective analysis 
```

#### Production curve

```{r, fig.align='center'}
## Production curve
plotspict.production(res_2) # check that the points fall around the curve
```

#### Model robustness

```{r}
## Model robustness (change of initial values)
set.seed(1234)
a_2=check.ini(inp_2, ntrials=30 
            ,verbose = FALSE)
#a_2$check.ini$propchng
table(a_2$check.ini$resmat[1:30] == 0) # check distance is close to 0 
```

#### Confidence intervals

```{r}
# Confidence intervals
fit_2=res_2 # no more than 1 unit
get.par("logBmBmsy", fit_2, exp=TRUE, CI = 0.9) 
get.par("logFmFmsy", fit_2, exp=TRUE, CI = 0.9)
```

#### Estimated biomass vs Simulated biomass

```{r}
# get.par("logB", fit_2, exp=TRUE, CI = 0.9) 
biomass_SPiCT_2 <- get.par("logB", res_2, exp = TRUE)[seq(1,length(inp_2$time),by=16),2]
biomass_SPiCT_2 <- biomass_SPiCT_2[1:15] 
```

```{r, fig.align= TRUE}
plot(biomass_SPiCT_2,
     col = "red",
     type = "l",
     ylim = c(min(simulated_biomass), max(biomass_SPiCT_2))
     )
lines(simulated_biomass, col = "blue", type = "l")
```

```{r}
# Error measures
glm_b_rmse <- rmse(biomass_SPiCT_2, simulated_biomass)
glm_b_mae <- mae(biomass_SPiCT_2, simulated_biomass)
glm_b_mape <- MAPE(biomass_SPiCT_2, simulated_biomass)
glm_b_rmse # RMSE
glm_b_mae # MAE
glm_b_mape # MAPE
```

### GAM (frequentist)

#### Measures for convergence

```{r}
# Diagnostic ---------------------------
## Convergence 
res_3$opt$convergence # 0 convergence  
all(is.finite(res_3$sd)) # if it is TRUE is okay 
```

#### Residuals analysis

```{r, fig.align='center'}
residuals_3 <- calc.osa.resid(res_3) # residuals of the model
plotspict.diagnostic(residuals_3) # test for the diagnostic
```

#### Retrospective analysis

```{r, fig.align='center'}
## Retrospective analysis
r_3 <- fit.spict(inp_3) 
rep_3 = retro(r_3, nretroyear=3) # Eliminate 3 years
plotspict.retro(rep_3) # plot for the retrospective analysis 
```

#### Production curve

```{r, fig.align='center'}
## Production curve
plotspict.production(res_3) # check that the points fall around the curve
```

#### Model robustness

```{r}
## Model robustness (change of initial values)
set.seed(1234)
a_3=check.ini(inp_3, ntrials=30 
            ,verbose = FALSE)
#a_3$check.ini$propchng
table(a_3$check.ini$resmat[1:30] == 0) # check distance is close to 0 
```

#### Confidence intervals

```{r}
# Confidence intervals
fit_3=res_3 # no more than 1 unit
get.par("logBmBmsy", fit_3, exp=TRUE, CI = 0.9) 
get.par("logFmFmsy", fit_3, exp=TRUE, CI = 0.9)
```

#### Estimated biomass vs Simulated biomass

```{r}
# get.par("logB", fit_3, exp=TRUE, CI = 0.9) 
biomass_SPiCT_3 <- get.par("logB", res_3, exp = TRUE)[seq(1,length(inp_3$time),by=16),2]
biomass_SPiCT_3 <- biomass_SPiCT_3[1:15] 
```

```{r, fig.align= TRUE}
plot(biomass_SPiCT_3,
     col = "red",
     type = "l",
     ylim = c(min(simulated_biomass), max(biomass_SPiCT_3))
     )
lines(simulated_biomass, col = "blue", type = "l")
```

```{r}
# Error measures
gam_f_rmse <- rmse(biomass_SPiCT_3, simulated_biomass)
gam_f_mae <- mae(biomass_SPiCT_3, simulated_biomass)
gam_f_mape <- MAPE(biomass_SPiCT_3, simulated_biomass)
gam_f_rmse # RMSE
gam_f_mae # MAE
gam_f_mape # MAPE
```

### GAM (Bayesian)

#### Measures for convergence

```{r}
# Diagnostic ---------------------------
## Convergence 
res_4$opt$convergence # 0 convergence  
all(is.finite(res_4$sd)) # if it is TRUE is okay 
```

#### Residuals analysis

```{r, fig.align='center'}
residuals_4 <- calc.osa.resid(res_4) # residuals of the model
plotspict.diagnostic(residuals_4) # test for the diagnostic
```

#### Retrospective analysis

```{r, fig.align='center'}
## Retrospective analysis
r_4 <- fit.spict(inp_4) 
rep_4 = retro(r_4, nretroyear=3) # Eliminate 3 years
plotspict.retro(rep_4) # plot for the retrospective analysis 
```

#### Production curve

```{r, fig.align='center'}
## Production curve
plotspict.production(res_4) # check that the points fall around the curve
```

#### Model robustness

```{r}
## Model robustness (change of initial values)
set.seed(1234)
a_4=check.ini(inp_4, ntrials=30 
            ,verbose = FALSE)
#a_4$check.ini$propchng
table(a_4$check.ini$resmat[1:30] == 0) # check distance is close to 0 
```

#### Confidence intervals

```{r}
# Confidence intervals
fit_4=res_4 # no more than 1 unit
get.par("logBmBmsy", fit_4, exp=TRUE, CI = 0.9) 
get.par("logFmFmsy", fit_4, exp=TRUE, CI = 0.9)
```

#### Estimated biomass vs Simulated biomass

```{r}
# get.par("logB", fit_4, exp=TRUE, CI = 0.9) 
biomass_SPiCT_4 <- get.par("logB", res_4, exp = TRUE)[seq(1,length(inp_4$time),by=16),2]
biomass_SPiCT_4 <- biomass_SPiCT_4[1:15] 
```

```{r, fig.align= TRUE}
plot(biomass_SPiCT_4,
     col = "red",
     type = "l",
     ylim = c(min(simulated_biomass), max(biomass_SPiCT_4))
     )
lines(simulated_biomass, col = "blue", type = "l")
```

```{r}
# Error measures
gam_b_rmse <- rmse(biomass_SPiCT_4, simulated_biomass)
gam_b_mae <- mae(biomass_SPiCT_4, simulated_biomass)
gam_b_mape <- MAPE(biomass_SPiCT_4, simulated_biomass)
gam_b_rmse # RMSE
gam_b_mae # MAE
gam_b_mape # MAPE
```

### Geostatistic

#### Measures for convergence

```{r}
# Diagnostic ---------------------------
## Convergence 
res_5$opt$convergence # 0 convergence  
all(is.finite(res_5$sd)) # if it is TRUE is okay 
```

#### Residuals analysis

```{r, fig.align='center'}
residuals_5 <- calc.osa.resid(res_5) # residuals of the model
plotspict.diagnostic(residuals_5) # test for the diagnostic
```

#### Retrospective analysis

```{r, fig.align='center'}
## Retrospective analysis
r_5 <- fit.spict(inp_5) 
rep_5 = retro(r_5, nretroyear=3) # Eliminate 3 years
plotspict.retro(rep_5) # plot for the retrospective analysis 
```

#### Production curve

```{r, fig.align='center'}
## Production curve
plotspict.production(res_5) # check that the points fall around the curve
```

#### Model robustness

```{r}
## Model robustness (change of initial values)
set.seed(1234)
a_5=check.ini(inp_5, ntrials=30 
            ,verbose = FALSE)
#a_5$check.ini$propchng
table(a_5$check.ini$resmat[1:30] == 0) # check distance is close to 0 
```

#### Confidence intervals

```{r}
# Confidence intervals
fit_5=res_5 # no more than 1 unit
get.par("logBmBmsy", fit_5, exp=TRUE, CI = 0.9) 
get.par("logFmFmsy", fit_5, exp=TRUE, CI = 0.9)
```

#### Estimated biomass vs Simulated biomass

```{r}
# get.par("logB", fit_5, exp=TRUE, CI = 0.9) 
biomass_SPiCT_5 <- get.par("logB", res_5, exp = TRUE)[seq(1,length(inp_5$time),by=16),2]
biomass_SPiCT_5 <- biomass_SPiCT_5[1:15] 
```

```{r, fig.align= TRUE}
plot(biomass_SPiCT_5,
     col = "red",
     type = "l",
     ylim = c(min(simulated_biomass), max(biomass_SPiCT_5))
     )
lines(simulated_biomass, col = "blue", type = "l")
```

```{r}
# Error measures
geo_b_rmse <- rmse(biomass_SPiCT_5, simulated_biomass)
geo_b_mae <- mae(biomass_SPiCT_5, simulated_biomass)
geo_b_mape <- MAPE(biomass_SPiCT_5, simulated_biomass)
geo_b_rmse # RMSE
geo_b_mae # MAE
geo_b_mape # MAPE
```

### Marked point pattern

#### Measures for convergence

```{r}
# Diagnostic ---------------------------
## Convergence 
res_6$opt$convergence # 0 convergence  
all(is.finite(res_6$sd)) # if it is TRUE is okay 
```

#### Residuals analysis

```{r, fig.align='center'}
residuals_6 <- calc.osa.resid(res_6) # residuals of the model
plotspict.diagnostic(residuals_6) # test for the diagnostic
```

#### Retrospective analysis

```{r, fig.align='center'}
## Retrospective analysis
r_6 <- fit.spict(inp_6) 
rep_6 = retro(r_6, nretroyear=3) # Eliminate 3 years
plotspict.retro(rep_6) # plot for the retrospective analysis 
```

#### Production curve

```{r, fig.align='center'}
## Production curve
plotspict.production(res_6) # check that the points fall around the curve
```

#### Model robustness

```{r}
## Model robustness (change of initial values)
set.seed(1234)
a_6=check.ini(inp_6, ntrials=30 
            ,verbose = FALSE)
#a_6$check.ini$propchng
table(a_6$check.ini$resmat[1:30] == 0) # check distance is close to 0 
```

#### Confidence intervals

```{r}
# Confidence intervals
fit_6=res_6 # no more than 1 unit
get.par("logBmBmsy", fit_6, exp=TRUE, CI = 0.9) 
get.par("logFmFmsy", fit_6, exp=TRUE, CI = 0.9)
```

#### Estimated biomass vs Simulated biomass

```{r}
# get.par("logB", fit_5, exp=TRUE, CI = 0.9) 
biomass_SPiCT_6 <- get.par("logB", res_6, exp = TRUE)[seq(1,length(inp_6$time),by=16),2]
biomass_SPiCT_6 <- biomass_SPiCT_6[1:15] 
```

```{r, fig.align= TRUE}
plot(biomass_SPiCT_6,
     col = "red",
     type = "l",
     ylim = c(min(simulated_biomass), max(biomass_SPiCT_6))
     )
lines(simulated_biomass, col = "blue", type = "l")
```

```{r}
# Error measures
geo_pref_b_rmse <- rmse(biomass_SPiCT_6, simulated_biomass)
geo_pref_b_mae <- mae(biomass_SPiCT_6, simulated_biomass)
geo_pref_b_mape <- MAPE(biomass_SPiCT_6, simulated_biomass)
geo_pref_b_rmse # RMSE
geo_pref_b_mae # MAE
geo_pref_b_mape # MAPE
```

```{r, include = FALSE}
data_error_pref <- data.frame(models = c("GLM F",
                                         "GLM B",
                                         "GAM F",
                                         "GAM B",
                                         "GEO",
                                         "MARKED PP"),
                            
                             rmse = c(glm_f_rmse,
                                      glm_b_rmse,
                                      gam_f_rmse,
                                      gam_b_rmse,
                                      geo_b_rmse,
                                      geo_pref_b_rmse),
                             
                             mae = c(glm_f_mae,
                                     glm_b_mae,
                                     gam_f_mae,
                                     gam_b_mae,
                                     geo_b_mae,
                                     geo_pref_b_mae),
                             
                             mape = c(glm_f_mape,
                                      glm_b_mape,
                                      gam_f_mape,
                                      gam_b_mape,
                                      geo_b_mape,
                                      geo_pref_b_mape))
```

# Error measures 

## Relativa biomass indices: random sampling

```{r, echo = FALSE}
knitr::kable(data_error_random, col.names = c("MODELS","RMSE", "MAE", "MAPE"), align = 'c')
```

## CPUE indices: preferential sampling

```{r, echo = FALSE}
knitr::kable(data_error_pref, col.names = c("MODELS","RMSE", "MAE", "MAPE"), align = 'c')
```
